---
title: "Area of the Ocean that is Fished"
subtitle: "The Global Footprint of Fisheries"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, comment = FALSE, message = FALSE)
```


```{r}
library(bigrquery)
library(dplyr)
library(raster)
library(sp)
library(rgdal)
library(sf)
library(rgeos)
library(ggplot2)
```
```{r, echo=FALSE}
BQ_connection <-  dbConnect(dbi_driver(), dataset = "", project = "world-fishing-827")

```



```{sql connection = BQ_connection, output.var = "total_fishing_effort_05"}
SELECT
  SUM(fishing_hours)/(COS(RADIANS(lat_bin/2)) * POW(111, 2)) hours_sq_km,
  lat_bin,
  lon_bin
FROM (
  SELECT
    FLOOR((lat_bin/100)/0.5) * 0.5  lat_bin,
    FLOOR((lon_bin/100)/0.5) * 0.5  lon_bin,
    fishing_hours
  FROM
   // [global-fishing-watch:global_footprint_of_fisheries.fishing_effort]
   [world-fishing-827:gfw_draft_data.fishing_effort]
  WHERE
    _PARTITIONTIME >= "2016-01-01 00:00:00"
    AND _PARTITIONTIME < "2017-01-01 00:00:00"
    AND fishing_hours > 0)
GROUP BY
  lat_bin,
  lon_bin
```

Generate a grid of points at 0.5 
```{r}
grid_latlon <- expand.grid(lat =seq(min(total_fishing_effort_05$lat_bin),
                                    max(total_fishing_effort_05$lat_bin), 0.5), 
                           lon = seq(min(total_fishing_effort_05$lon_bin),
                                     max(total_fishing_effort_05$lon_bin), 0.5))
```

Join the global grid to the global fishing data. Assign locations with no fishing
a fishing value of 0.
```{r}
gridded_fishing = left_join(grid_latlon, total_fishing_effort_05, by = c('lat' = 'lat_bin', 'lon' = 'lon_bin'))
gridded_fishing$hours = ifelse(is.na(gridded_fishing$hours), 0, gridded_fishing$hours)
```

Convert fishing effort dataframe into a spatial dataframe (specify lon and lat as x, y)
```{r}

coordinates(gridded_fishing)=~lon+lat
```

Assign a projection to the gridded fishing effort
```{r}
# set it to lat-long
proj4string(gridded_fishing)=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") 

#ensure fishing in on regular grid
gridded(gridded_fishing) = TRUE
```


Generate the fishing effort raster. Specify original projection and extent, 
then project the raster into Gall-Peters equal area
```{r}
fishing_raster_05 = rasterFromXYZ(gridded_fishing, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
bb = extent(-180.25, 179.75, -88.25, 81.75)
fishing_raster_05 = setExtent(fishing_raster_05, bb, keepres = TRUE)
#this is Gall-Peters Equal Area Projection
fishing_raster_05 = projectRaster(fishing_raster_05, crs = "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +datum=WGS84 +units=m +no_defs", over = TRUE) 
```

Export raster 
```{r, eval=FALSE}
# writeRaster(fishing_raster_05, '~/Documents/GlobalFishingWatch/Global_raster_paper/oceans_equal_area/fishing_raster_05_20180228.tif', format = 'GTiff', overwrite = TRUE)
writeRaster(fishing_raster_05, '../../data/fishing_raster_05_20180228.tif', format = 'GTiff', overwrite = TRUE)
```


Either read in fishing raster of continue from above.
Load an oceans shapefile 
```{r}
#fishing_raster_05 <- raster('~/Documents/GlobalFishingWatch/Projects/Global_raster_paper/oceans_equal_area/fishing_raster_05.tif')
ocean_cea_sf <- read_sf('../../data/oceans_cea/oceans_cea.shp')
```

Reproject the oceans shapefile using a Gall-Peters Equal Area Projection to ensure equal area at all latitudes.
```{r}
ocean_cea_sf <- st_transform(ocean_cea_sf, crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
```

Check the vector file and projection look appropriate
```{r}
head(ocean_cea_sf)
ggplot() +
    geom_sf(data = ocean_cea_sf, fill = 'dodgerblue')
```

For simplicity areas of fishing a value of 1 and those without NA.
```{r}
fishing_raster_05[fishing_raster_05[] > 0] <- 1 
fishing_raster_05[fishing_raster_05[] <= 0] <- NA 
```

A quick visual check to see if the raster looks appropriate (though it will not yet have the appropriate projection)
```{r}
plot(fishing_raster_05)
```

Convert the raster to a vector which makes it possible to compare with the oceans vector.

```{r, cache=TRUE}
#raster to polygon using maximum number of vertices (n = 16)
fishing_polygon <- rasterToPolygons(fishing_raster_05, dissolve = TRUE, n = 16)
#convert Spatial dataframe to Simple features
fishing_polygon_sf <- st_as_sf(fishing_polygon)
#reproject fishing to Gall-Peters Equal area projection
fishing_polygon_sf <- st_transform(fishing_polygon_sf, crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
```


Identify where the fishing and ocean vector files intersect. Ensures we clip the
fishing raster to the ocean.
```{r, cache=TRUE}
fishing_ocean_intersect <- sf::st_intersection(x = ocean_cea_sf, fishing_polygon_sf)
```

Calculate the area of the fishing/ocean intersection vector and the area of the ocean vector, both of which are in equal area projection. The units are relatively unimportant, but in m^2^. Calculate the percent of the ocean that is covered fishing.
```{r}
fishing_area <- st_area(fishing_ocean_intersect)
ocean_area <- st_area(ocean_cea_sf)
(fishing_area/ocean_area) * 100
```

```{r}
library(ggplot2)
ggplot() +
    geom_sf(data = fishing_ocean_intersect, fill = 'dodgerblue')
```


#### Same analysis using previous fishing raster

```{r}
fishing_raster_05_old <- raster('~/Documents/GlobalFishingWatch/Projects/Global_raster_paper/oceans_equal_area/fishing_raster_05.tif')
```

For simplicity areas of fishing a value of 1 and those without NA.
```{r}
fishing_raster_05_old[fishing_raster_05_old[] > 0] <- 1 
fishing_raster_05_old[fishing_raster_05_old[] <= 0] <- NA 
```

A quick visual check to see if the raster looks appropriate (though it will not yet have the appropriate projection)
```{r}
plot(fishing_raster_05_old)
```

Convert the raster to a vector which makes it possible to compare with the oceans vector.

```{r, cache=TRUE}
#raster to polygon using maximum number of vertices (n = 16)
fishing_polygon_old <- rasterToPolygons(fishing_raster_05_old, dissolve = TRUE, n = 16)
#convert Spatial dataframe to Simple features
fishing_polygon_old_sf <- st_as_sf(fishing_polygon_old)
#reproject fishing to Gall-Peters Equal area projection
fishing_polygon_old_sf <- st_transform(fishing_polygon_old_sf, crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
```


Identify where the fishing and ocean vector files intersect. Ensures we clip the
fishing raster to the ocean.
```{r, cache=TRUE}
fishing_old_ocean_intersect <- sf::st_intersection(x = ocean_cea_sf, fishing_polygon_old_sf)
```

Calculate the area of the fishing/ocean intersection vector and the area of the ocean vector, both of which are in equal area projection. The units are relatively unimportant, but in m^2^. Calculate the percent of the ocean that is covered fishing.
```{r}
fishing_area_old <- st_area(fishing_old_ocean_intersect)
ocean_area <- st_area(ocean_cea_sf)
(fishing_area_old/ocean_area) * 100
```

