---
title: "Area of the Ocean that is Fished"
subtitle: "The Global Footprint of Fisheries"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, comment = FALSE, message = FALSE)
```

Load packages
```{r}
library(bigrquery)
library(dplyr)
library(raster)
library(sp)
library(rgdal)
library(sf)
library(rgeos)
library(ggplot2)
```

```{r, echo=FALSE}
BQ_connection <-  dbConnect(dbi_driver(), dataset = "", project = "world-fishing-827")

```


Query to get the fishing effort at 0.5 degree resolution from the GFW public Big Query
table. [Fishing Effort](https://bigquery.cloud.google.com/table/global-fishing-watch:global_footprint_of_fisheries.fishing_effort?pli=1)
```{sql connection = BQ_connection, output.var = "total_fishing_effort_05"}
SELECT
  SUM(fishing_hours)/(COS(RADIANS(lat_bin/2)) * POW(111, 2)) hours_sq_km,
  lat_bin,
  lon_bin
FROM (
  SELECT
    FLOOR((lat_bin/100)/0.5) * 0.5  lat_bin,
    FLOOR((lon_bin/100)/0.5) * 0.5  lon_bin,
    fishing_hours
  FROM
    [global-fishing-watch:global_footprint_of_fisheries.fishing_effort]
  WHERE
    _PARTITIONTIME >= "2016-01-01 00:00:00"
    AND _PARTITIONTIME < "2017-01-01 00:00:00"
    AND fishing_hours > 0)
GROUP BY
  lat_bin,
  lon_bin
```

Generate a grid of points at 0.5 
```{r}
grid_latlon <- expand.grid(lat =seq(min(total_fishing_effort_05$lat_bin),
                                    max(total_fishing_effort_05$lat_bin), 0.5), 
                           lon = seq(min(total_fishing_effort_05$lon_bin),
                                     max(total_fishing_effort_05$lon_bin), 0.5))
```

Join the global grid to the global fishing data. Assign locations with no fishing
a fishing value of 0.
```{r}
gridded_fishing = left_join(grid_latlon, total_fishing_effort_05, by = c('lat' = 'lat_bin', 'lon' = 'lon_bin'))
gridded_fishing$hours = ifelse(is.na(gridded_fishing$hours), 0, gridded_fishing$hours)
```

Convert fishing effort dataframe into a spatial dataframe (specify lon and lat as x, y)
```{r}
coordinates(gridded_fishing)=~lon+lat
```

Specify `EPSG:4326` as initial projection for the gridded fishing effort. 
```{r}
# set it to EPSG:4326
proj4string(gridded_fishing)=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") 

#ensure fishing in on regular grid
gridded(gridded_fishing) = TRUE
```


Generate the fishing effort raster. Specify original projection and extent, 
then project the raster into Gall-Peters equal area projection
```{r}
fishing_raster_05 = rasterFromXYZ(gridded_fishing, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
bb = extent(-180.25, 179.75, -88.25, 81.75)
fishing_raster_05 = setExtent(fishing_raster_05, bb, keepres = TRUE)
#this is Gall-Peters Equal Area Projection
fishing_raster_05 = projectRaster(fishing_raster_05, crs = "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +datum=WGS84 +units=m +no_defs", over = TRUE) 
```

Export raster 
```{r, eval=FALSE}
writeRaster(fishing_raster_05, '~/Documents/GlobalFishingWatch/Global_raster_paper/oceans_equal_area/fishing_raster_05_20180228.tif', format = 'GTiff', overwrite = TRUE)
```


Load a 10m Oceans shapefile from Natural Earth http://www.naturalearthdata.com/downloads/10m-physical-vectors/:  
This file has had polygons dissolved into single ocean polygon and has been projected into Gall-Peters Equal area projections already. 
```{r}
ocean_cea_sf <- read_sf('~/Documents/GlobalFishingWatch/shapefiles/oceans_CEA_equal_area/oceans_cea.shp')
```

For completeness and to ensure the appropriate projection, here we reproject the oceans 
shapefile using a Gall-Peters Equal Area Projection to ensure equal area at all latitudes.
```{r}
ocean_cea_sf <- st_transform(ocean_cea_sf, crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
```

Check the vector file to ensure projection look appropriate
```{r}
head(ocean_cea_sf)
ggplot() +
    geom_sf(data = ocean_cea_sf, fill = 'dodgerblue')
```

For simplicity assign areas of fishing a value of 1 and those without NA.
```{r}
fishing_raster_05[fishing_raster_05[] > 0] <- 1 
fishing_raster_05[fishing_raster_05[] <= 0] <- NA 
```

A quick visual check to see if the raster looks appropriate (though it will not yet have the appropriate projection)
```{r}
plot(fishing_raster_05)
```

Convert the raster to a vector file which makes it possible to compare with the oceans vector.
```{r}
#raster to polygon using maximum number of vertices (n = 16)
fishing_polygon <- rasterToPolygons(fishing_raster_05, dissolve = TRUE, n = 16)
#convert Spatial dataframe to Simple features
fishing_polygon_sf <- st_as_sf(fishing_polygon)
#reproject fishing to Gall-Peters Equal area projection
fishing_polygon_sf <- st_transform(fishing_polygon_sf, crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
```


Identify where the fishing and ocean vector files intersect. Ensures that for grid cells 
that overlap the coastline we only include the area that is within the ocean. 
In some sense, this provides a smaller grid along the coastline.
```{r}
fishing_ocean_intersect <- sf::st_intersection(x = ocean_cea_sf, fishing_polygon_sf)
```

Calculate the area of the fishing/ocean intersection vector and the area of the ocean vector, both of which are in equal area projection. The units are relatively unimportant, but in m^2^. Calculate the percent of the ocean that is covered fishing.
```{r}
fishing_area <- st_area(fishing_ocean_intersect)
ocean_area <- st_area(ocean_cea_sf)
(fishing_area/ocean_area) * 100
```

```{r}
library(ggplot2)
ggplot() +
    geom_sf(data = fishing_ocean_intersect, fill = 'dodgerblue')
```


#### Same analysis using previous fishing raster
Here we illustrate the same calculation using a fishing raster that
was generated using an earlier version of the fishing effort dataset. This dataset
may have overemphasized squid-jigger fishing.
```{r}
fishing_raster_05_old <- raster('~/Documents/GlobalFishingWatch/Projects/Global_raster_paper/oceans_equal_area/fishing_raster_05.tif')
```

Analysis same as before: Convert grid cells with fishing to 1's and grid cells without fishing to NA.
```{r}
fishing_raster_05_old[fishing_raster_05_old[] > 0] <- 1 
fishing_raster_05_old[fishing_raster_05_old[] <= 0] <- NA 
```

A quick visual check to see if the raster looks appropriate (though it will not yet have the appropriate projection).
```{r}
plot(fishing_raster_05_old)
```

Convert the raster to a vector, which makes it possible to compare with the oceans vector.
```{r}
#raster to polygon using maximum number of vertices (n = 16)
fishing_polygon_old <- rasterToPolygons(fishing_raster_05_old, dissolve = TRUE, n = 16)
#convert Spatial dataframe to Simple features
fishing_polygon_old_sf <- st_as_sf(fishing_polygon_old)
#reproject fishing to Gall-Peters Equal area projection
fishing_polygon_old_sf <- st_transform(fishing_polygon_old_sf, crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
```


Identify where the fishing and ocean vector files intersect. Ensures we clip the
fishing raster to the ocean.
```{r}
fishing_old_ocean_intersect <- sf::st_intersection(x = ocean_cea_sf, fishing_polygon_old_sf)
```

Calculate the area of the fishing/ocean intersection vector and the area of the ocean vector, both of which are in equal area projection. The units are relatively unimportant, but in m^2^. Calculate the percent of the ocean that is covered fishing.
```{r}
fishing_area_old <- st_area(fishing_old_ocean_intersect)
ocean_area <- st_area(ocean_cea_sf)
(fishing_area_old/ocean_area) * 100
```


Note that this value is ~55%
