---
title: "Fishing Coverage- Equal Area Projection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo = FALSE, message=FALSE, comment=FALSE}
library(tidyverse)
library(viridis)
library(extrafont)
library(bigrquery)
library(rgdal)
library(rgeos)
library(ggalt)
library(printr)
loadfonts()
BQ_connection <-  dbConnect(dbi_driver(), dataset = "", project = "world-fishing-827")
```

This notebook calculates the area of the ocean over which fishing by all gear 
types takes place, as well as fishing by the three (3) primary geartypes described
in the manuscript (drifting longlines, trawlers, purse seines).

The code starts with an example using all fishing and is then followed by
individual analyses for each gear type.

Query to generate total fishing raster at 0.5 degree resolution.
```{sql connection = BQ_connection, output.var = "total_fishing_effort_05"}
SELECT
 SUM(hours) hours,
 FLOOR(lat/0.5)*0.5 lat_bin,
 FLOOR(lon/0.5)*0.5 lon_bin
FROM (
 SELECT
   *
 FROM (
   SELECT
     mmsi,
     hours,
     seg_id,
     lat,
     lon
   FROM
    [gfw_research.nn]
    /*[gfw_research.FAO] */
   WHERE
   _PARTITIONTIME BETWEEN TIMESTAMP('2016-01-01') AND TIMESTAMP('2016-12-31')
   AND
     nnet_score >.5 ) /*measure_new_score */
 WHERE
   seg_id IN (
    SELECT seg_id FROM [world-fishing-827:gfw_research.segments2]
        WHERE point_count > 20 
        AND max_lat - min_lat > .01 
        AND max_lon - min_lon > .01
        AND NOT (point_count<10
    AND terrestrial_positions = point_count)
  AND NOT ((min_lon >= 0 /* these are almost definitely noise*/
      AND max_lon <= 0.109225)
    OR (min_lat >= 0
      AND max_lat <= 0.109225) OR
      (min_lat <= -0.109225 AND
      max_lat <= 0))) /*must move 100 meters in lat and lon*/
AND mmsi IN (SELECT mmsi from [gfw_research.vessel_info_20170522] WHERE on_fishing_list_nn) 
AND mmsi not in (SELECT mmsi from [world-fishing-827:gfw_research.vessel_info_20170522] WHERE offsetting)
AND mmsi not in (725000332,
    725000910,
    725000335,
    725000410) /*trawlers off chile classified as PS*/
AND mmsi not in (224084620, 224107670) /* trawlers South Atlantic as PS*/
AND mmsi not in (431560000, 431757000, 432846000) /*Western Pacific research as PS*/
AND mmsi not in (431704470, 431267000, 431800050) /*north west pacific, squid jiggers as PS, and research vessel FRA */
AND mmsi not in (416000002) /* south atlantic */
ANd mmsi not in (431700260, 431797000,431704490) /* offsetting longliners, Panama */
and mmsi not in (412421007) /* DL track in EEZ north Pacific */
And mmsi not in (413322690,413322770) /*DL track in EEZ near NZ, good longliners, odd fishing */
AND mmsi not in (412420971, 412420973) /* odd squid jiggers tracking across Pacific */
AND mmsi not in (416121800) /*south Pacific track across EEZ */
AND mmsi NOT IN (412421005) /*squid jigger appears to be fishing across southern Atlantic during transit */
AND mmsi NOT IN (441051000,412420001) /*trawler, transitting south Atlantic */
AND mmsi NOT IN (432288000) /* research vessel, track across West Pacific EEZ, listed as drifting longliner */ ) a
INNER JOIN
 (SELECT
 mmsi
FROM
 [scratch_david_mmsi_lists.nn_labels_20170516]
where label IN ("fixed_gear",
   "drifting_longlines",
   "other_fishing",
   "trawlers",
   "purse_seines",
   "squid_jigger")
and mmsi in (SELECT mmsi from  [scratch_david_mmsi_lists.nn_labels_2016_20170516]
WHERE label IN ("fixed_gear",
   "drifting_longlines",
   "other_fishing",
   "trawlers",
   "purse_seines",
   "squid_jigger")
)) b
ON
 a.mmsi = b.mmsi
GROUP BY
 lat_bin,
 lon_bin)
```

Generate a grid of points at 0.5 
```{r}
#nrow(total_fishing_effort_05)

grid_latlon <- expand.grid(lat =seq(min(total_fishing_effort_05$lat_bin),
                                    max(total_fishing_effort_05$lat_bin), 0.5), 
                           lon = seq(min(total_fishing_effort_05$lon_bin),
                                     max(total_fishing_effort_05$lon_bin), 0.5))

```

Join the global grid to the global fishing data. Assign locations with no fishing
a fishing value of 0.
```{r}
gridded_fishing = left_join(grid_latlon, total_fishing_effort_05, by = c('lat' = 'lat_bin', 'lon' = 'lon_bin'))
gridded_fishing$hours = ifelse(is.na(gridded_fishing$hours), 0, gridded_fishing$hours)
#gridded_fishing %>% filter(hours == 0) %>% summarize(n())
#nrow(gridded_fishing) - gridded_fishing %>% filter(hours == 0) %>% summarize(n())
```

Convert fishing effort dataframe into a spatial dataframe (specify lon and lat as x, y)
```{r}
library(sp)
library(rgdal)
coordinates(gridded_fishing)=~lon+lat
```

Assign a projection to the gridded fishing effort
```{r}
proj4string(gridded_fishing)=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") # set it to lat-long
#gridded_fishing = spTransform(gridded_fishing,CRS("+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +datum=WGS84 +units=m +no_defs"))

#ensure fishing in on regular grid
gridded(gridded_fishing) = TRUE
```


Generate the fishing effort raster. Specify original projection and extent, 
then project the raster into Gall-Peters equal area
```{r}
library(raster)
fishing_raster_05 = rasterFromXYZ(gridded_fishing, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
bb = extent(-180, 180, -78.8, 81.8)
fishing_raster_05 = setExtent(fishing_raster_05, bb, keepres = TRUE)
fishing_raster_05 = projectRaster(fishing_raster_05, crs = "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +datum=WGS84 +units=m +no_defs", over = TRUE) #this is Gall-Peters Equal Area Projection
```


Export raster which is then analyzed in QGIS (See outline below).
```{r, eval=FALSE}
writeRaster(fishing_raster_05, '~/Documents/GlobalFishingWatch/Global_raster_paper/oceans_equal_area/fishing_raster_05_round2.tif', format = 'GTiff', overwrite = TRUE)
```

### Loading the oceans shapefile
```{r}
setwd("~/Documents/GlobalFishingWatch/oceans_equal_area/")
oceans <- readOGR('.', "oceans_gallpeters")
oceans_df = fortify(oceans)
```




Building upon this post:
https://gis.stackexchange.com/questions/147158/how-to-calculate-the-area-of-a-raster-within-a-polygon-which-is-greater-than-a-s

Then in QGIS:
*1 Import both the raster and the gall_peters projected oceans shapefile (oceans
shapefile from natural earth http://www.naturalearthdata.com/downloads/10m-physical-vectors/).
*2 Within raster tools, use Raster Calculator to calculate where the fishing value is greater than 0. 
fishing_raster_05@1 > 0, this is a Boolean that gives 1 if true, 0 if false and generates a new raster of 0/1.
*3 Then use the Zonal Statistics (under Raster) to generate the mean value for this 0/1 raster within the oceans polygon.
* This represents the percentage of the ocean that is fished. Knowing the areas of the polygon one could calculate the actual area fished.

For the total effort raster = 56.4%

## For each gear type % of total ocean (% of total area fished)
These data arise from analysis below

longline = 45.7% or (81%)
trawler = 9.7% or (17.2%)
purse = 17.1% or (30.3%)


Using the field calculator and the Gall-Peters projected oceans shapefile I calculated an area of the shapefile
Attributes Table -> Field Calculator -> area($geometry) in square meters. 363223868892795 m2
Converting to square kilometers the value agrees with values reported online: #363223868 km2

Total area fished = 0.56 * 363223868 = 203,405,366 km2 or 2.03 x 10^8 km2


```{r}
# % of whole ocean
total_fishing_percent = 0.564
longliner_percent = 0.457
trawler_percent = 0.097
purse_seine_percent = 0.171


#for the total ocean
total_ocean_area_km2 = 363223868

total_fished_area = total_fishing_percent * total_ocean_area_km2
longliner_ocean = longliner_percent * total_ocean_area_km2
trawler_ocean = trawler_percent * total_ocean_area_km2
purse_seine_ocean = purse_seine_percent * total_ocean_area_km2


df = data.frame(gear_type = c('total','longliner', 'trawler', 'purse_seine'),
                area_of_ocean_fished = c(total_fished_area, longliner_ocean, trawler_ocean, purse_seine_ocean))
df

```

ANALYSIS/DATA for individual gear types following methods outlines in detail
above.

### Run the query to get the fishing effort by type so that individual rasters 
can be created for each gear type

```{sql connection = BQ_connection, output.var = "fishing_effort_by_type_05"}
SELECT
 SUM(hours) hours,
 FLOOR(lat/0.5)*0.5 lat_bin,
 FLOOR(lon/0.5)*0.5 lon_bin,
 label
FROM (
 SELECT
   *
 FROM (
   SELECT
     mmsi,
     hours,
     seg_id,
     lat,
     lon
   FROM
    [gfw_research.nn]
   WHERE
   _PARTITIONTIME BETWEEN TIMESTAMP('2016-01-01') AND TIMESTAMP('2016-12-31')
   AND
     nnet_score >.5 )
 WHERE
seg_id IN (
    SELECT seg_id FROM [world-fishing-827:gfw_research.segments2]
        WHERE point_count > 20 
        AND max_lat - min_lat > .01 
        AND max_lon - min_lon > .01
        AND NOT (point_count<10
    AND terrestrial_positions = point_count)
  AND NOT ((min_lon >= 0 /* these are almost definitely noise*/
      AND max_lon <= 0.109225)
    OR (min_lat >= 0
      AND max_lat <= 0.109225) OR
      (min_lat <= -0.109225 AND
      max_lat <= 0))) /*must move 100 meters in lat and lon*/
AND mmsi IN (SELECT mmsi from [gfw_research.vessel_info_20170522] WHERE on_fishing_list_nn) 
AND mmsi not in (SELECT mmsi from [world-fishing-827:gfw_research.vessel_info_20170522] WHERE offsetting)
AND mmsi not in (725000332,
    725000910,
    725000335,
    725000410) /*trawlers off chile classified as PS*/
AND mmsi not in (224084620, 224107670) /* trawlers South Atlantic as PS*/
AND mmsi not in (431560000, 431757000, 432846000) /*Western Pacific research as PS*/
AND mmsi not in (431704470, 431267000, 431800050) /*north west pacific, squid jiggers as PS, and research vessel FRA */
AND mmsi not in (416000002) /* south atlantic */
ANd mmsi not in (431700260, 431797000,431704490) /* offsetting longliners, Panama */
and mmsi not in (412421007) /* DL track in EEZ north Pacific */
And mmsi not in (413322690,413322770) /*DL track in EEZ near NZ, good longliners, odd fishing */
AND mmsi not in (412420971, 412420973) /* odd squid jiggers tracking across Pacific */
AND mmsi not in (416121800) /*south Pacific track across EEZ */
AND mmsi NOT IN (760000650) /*oil products tanker, off Peru, labeled as trawler */
AND mmsi NOT IN (412421005) /*squid jigger appears to be fishing across southern Atlantic during transit */
AND mmsi NOT IN (413201110) /*longliner, currently classified as a trawler (mid-Atlantic) */
AND mmsi NOT IN (432288000) /* research vessel, track across West Pacific EEZ, listed as drifting longliner */
/*mmsi NOT IN (432621000,431028000,431494000, 432850000,9102213,112420884,150200144,150200148,150200162,150200250,150200252,150200389,150200783,150200893,150200895,150200974,150201019,150201073,150201423,150201498,150201517,150201536,150201550,150201552,150201776,150202857,150202858,150202904,150203018,150203019,150203020,150203021,150203022,158888888,200000590,200000596,200000598,200000599,411041802,412071033,412200146,412200991,412207532,412207533,412207535,412207536,412207537,412207538,412207539,412207541,412207542,412207545,412207546,412209011,412209092,412209093,412209097,412209098,412209099,412209101,412209102,412209105,412209106,412209132,412209133,412209136,412209139,412209141,412209168,412209171,412209172,412211196,412211691,412322329,412322547,412322587,412322866,412322952,412322953,412322984,412323207,412326016,412327269,412327285,412327422,412327493,412328279,412328794,412329642,412329643,412329644,412331106,412331107,412331109,412331111,412331112,412333121,412364066,412364739,412401220,412402610,412411514,412411515,412411519,412411529,412412981,412418059,412419955,412419958,412420248,412420276,412420421,412420453,412420502,412420503,412420561,412420573,412420574,412420575,412420576,412420684,412420787,412420789,412420805,412420833,412420871,412420878,412420879,412420881,412420882,412420883,412420885,412420886,412425706,412437961,412437962,412440077,412440078,412440104,412440105,412440106,412440113,412440255,412440256,412440257,412440258,412440259,412440261,412440377,412447093,412480433,412480434,412661896,412693030,412699340,413035015,413066568,413322650,413685698,414203586,415000005,441301074,462801187,800011282,800016350,800017747,800017800,800018712,800019446,800030681,812411524,900007134,900007138,900007140,900007148,900007154,900023867,900025166,900025206,900025236,900025344,900025350,900025357,900025393,900025395,910060382)  vessels with offset positions
AND mmsi NOT IN (SELECT mmsi FROM [world-fishing-827:scratch_david_mmsi_lists.research_20170522] research/training vessels)
and mmsi not in (987357573,987357579,987357559,986737000,983712160,987357529,987357559,987357579)  helicopters */) a
INNER JOIN
 (SELECT
 mmsi,
 label
FROM
 [scratch_david_mmsi_lists.nn_labels_20170516]
where label IN ("fixed_gear",
   "drifting_longlines",
   "other_fishing",
   "trawlers",
   "purse_seines",
   "squid_jigger")
and mmsi in (SELECT mmsi from  [scratch_david_mmsi_lists.nn_labels_2016_20170516]
WHERE label IN ("fixed_gear",
   "drifting_longlines",
   "other_fishing",
   "trawlers",
   "purse_seines",
   "squid_jigger")
)) b
ON
 a.mmsi = b.mmsi
GROUP BY
 lat_bin,
 lon_bin,
label)
```

Separate out the different gear types

```{r}
purse_seine_effort_05 = fishing_effort_by_type_05 %>% 
  filter(inferred_label == 'purse_seines'& lat_bin > -90 & lat_bin < 90)


longline_effort_05 = fishing_effort_by_type_05 %>%
    filter(inferred_label == 'drifting_longlines'& lat_bin > -90 & lat_bin < 85)

trawler_effort_05 = fishing_effort_by_type_05 %>%
    filter(inferred_label == 'trawlers' & lat_bin > -90 & lat_bin < 90) 
```


### Longliners
Apply the gridded lat lon and replace the NA values with 0

```{r}
gridded_fishing_longliner = left_join(grid_latlon, longline_effort_05, by = c('lat' = 'lat_bin', 'lon' = 'lon_bin'))
gridded_fishing_longliner$hours = ifelse(is.na(gridded_fishing_longliner$hours), 0, gridded_fishing_longliner$hours)
```

Same as above, generate a spatial dataframe and then a raster with the Gall-Peters projections
```{r}
coordinates(gridded_fishing_longliner)=~lon+lat
```



```{r}
gridded_raster_longliner = rasterFromXYZ(gridded_fishing_longliner, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") #specify original projections as longlat
bb = extent(-180, 180, -90, 90) #specify the extent
gridded_raster_longliner = setExtent(gridded_raster_longliner, bb, keepres = TRUE) #apply extent
gridded_raster_longliner = projectRaster(gridded_raster_longliner, crs = "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +datum=WGS84 +units=m +no_defs", over = TRUE) #project the raster into the Gall-Peters Equal Area Projection
```


```{r, eval = FALSE}
writeRaster(gridded_raster_longliner, '~/Documents/GlobalFishingWatch/oceans_equal_area/fishing_raster_05_longliner.tif', format = 'GTiff', overwrite = TRUE)
```

### Trawlers

```{r}
gridded_fishing_trawler = left_join(grid_latlon, trawler_effort_05, by = c('lat' = 'lat_bin', 'lon' = 'lon_bin'))
gridded_fishing_trawler$hours = ifelse(is.na(gridded_fishing_trawler$hours), 0, gridded_fishing_trawler$hours)
```


```{r}
coordinates(gridded_fishing_trawler)=~lon+lat
```



```{r}
gridded_raster_trawler = rasterFromXYZ(gridded_fishing_trawler, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") #specify original projections as longlat
bb = extent(-180, 180, -90, 90) #specify the extent
gridded_raster_trawler = setExtent(gridded_raster_trawler, bb, keepres = TRUE) #apply extent
gridded_raster_trawler = projectRaster(gridded_raster_trawler, crs = "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +datum=WGS84 +units=m +no_defs", over = TRUE) #project the raster into the Gall-Peters Equal Area Projection
```


```{r, eval = FALSE}
writeRaster(gridded_raster_trawler, '~/Documents/GlobalFishingWatch/oceans_equal_area/fishing_raster_05_trawler.tif', format = 'GTiff', overwrite = TRUE)
```



### Purse seines

```{r}
gridded_fishing_purseseine = left_join(grid_latlon, purse_seine_effort_05, by = c('lat' = 'lat_bin', 'lon' = 'lon_bin'))
gridded_fishing_purseseine$hours = ifelse(is.na(gridded_fishing_purseseine$hours), 0, gridded_fishing_purseseine$hours)
```


```{r}
coordinates(gridded_fishing_purseseine)=~lon+lat
```



```{r}
gridded_raster_purseseine = rasterFromXYZ(gridded_fishing_purseseine, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") #specify original projections as longlat
bb = extent(-180, 180, -90, 90) #specify the extent
gridded_raster_purseseine = setExtent(gridded_raster_purseseine, bb, keepres = TRUE) #apply extent
gridded_raster_purseseine = projectRaster(gridded_raster_purseseine, crs = "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +datum=WGS84 +units=m +no_defs", over = TRUE) #project the raster into the Gall-Peters Equal Area Projection
```


```{r, eval=FALSE}
writeRaster(gridded_raster_purseseine, '~/Documents/GlobalFishingWatch/oceans_equal_area/fishing_raster_05_purseseine.tif', format = 'GTiff', overwrite = TRUE)
```

