---
title: "Tracking the Global Footprint of Fisheries: Figure 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = FALSE, warning = FALSE)
```

```{r echo = FALSE, message=FALSE, comment=FALSE}
library(tidyverse)
library(viridis)
library(bigrquery)
library(rgdal)
library(rgeos)
library(sf)
BQ_connection <-  dbConnect(dbi_driver(), dataset = "", project = "world-fishing-827")
```

Code for making Figure 1 from Tracking the Footprint of Global Fisheries paper from
public Big Query datasets provided by Global Fishing Watch.

##### Figure prep
```{r}
theme_gfw_paper <-  theme(text = element_text(family="Arial", face="bold", size=12),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.line.x = element_line(colour = "black", size = .5),
        axis.line.y = element_line(colour = "black", size = .5),
        legend.position = "bottom",
        axis.title.y = element_text(size = 12, margin=margin(0,15,0,0)),
        axis.title.x = element_text(size = 12, margin=margin(15,0,0,0)),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        plot.margin = unit(c(1,1,1,1), "cm")
        )
```


Load shapefile downloaded from Natural Earth (110m resolution)
```{r}
land_df <- read_sf('../data/ne_110m_land/ne_110m_land.shp', quiet = TRUE)
# This shapefile is available from http://www.naturalearthdata.com/downloads/110m-physical-vectors/110m-land/
```


#### SQL query for Total Fishing Effort (0.5 degree)
```{sql connection = BQ_connection, output.var = "total_fishing_effort_05"}
SELECT
  SUM(fishing_hours)/(COS(RADIANS(lat_bin/2)) * POW(111, 2)) hours_sq_km,
  lat_bin,
  lon_bin
FROM (
  SELECT
    FLOOR((lat_bin/100)/0.5) * 0.5 lat_bin,
    FLOOR((lon_bin/100)/0.5) * 0.5 lon_bin,
    fishing_hours
  FROM
    [global-fishing-watch:global_footprint_of_fisheries.fishing_effort]
  WHERE
    _PARTITIONTIME >= "2016-01-01 00:00:00"
    AND _PARTITIONTIME < "2017-01-01 00:00:00"
    AND fishing_hours > 0)
GROUP BY
  lat_bin,
  lon_bin
```



#### SQL Query for Fishing Effort by Geartype (0.5 degree)
```{sql connection = BQ_connection, output.var = "fishing_effort_by_type_05"}
SELECT
  SUM(fishing_hours)/(COS(RADIANS(lat_bin/2)) * POW(111, 2)) hours_sq_km,
  lat_bin,
  lon_bin,
  geartype
FROM (
  SELECT
    FLOOR((lat_bin/100)/0.5) * 0.5 lat_bin,
    FLOOR((lon_bin/100)/0.5) * 0.5 lon_bin,
    fishing_hours,
    geartype
  FROM
    [global-fishing-watch:global_footprint_of_fisheries.fishing_effort]
  WHERE
    _PARTITIONTIME >= "2016-01-01 00:00:00"
    AND _PARTITIONTIME < "2017-01-01 00:00:00"
    AND fishing_hours > 0)
GROUP BY
  lat_bin,
  lon_bin,
  geartype
```


```{r}
head(total_fishing_effort_05)
```


```{r}
head(fishing_effort_by_type_05)
```

#### Process data
Data processing for plotting total fishing effort
```{r}
total_fishing_effort_05$hours_sq_km = total_fishing_effort_05$hours_sq_km * 10000
total_fishing_effort_05$logHours = log10(total_fishing_effort_05$hours_sq_km)
total_fishing_effort_05$logHours = ifelse(total_fishing_effort_05$logHours >= 5, 
                                          5, 
                                          total_fishing_effort_05$logHours)
total_fishing_effort_05$logHours = ifelse(total_fishing_effort_05$logHours <= 1, 
                                          1, 
                                          total_fishing_effort_05$logHours)

total_fishing_effort_05 = total_fishing_effort_05 %>%
    filter(lat_bin < 85 & lat_bin > -85 & lon_bin > -180 & lon_bin < 180)
```


Data processing for plotting fishing effort by geartype
```{r}

fishing_effort_by_type_05$hours_sq_km = fishing_effort_by_type_05$hours_sq_km * 10000
fishing_effort_by_type_05$logHours = log10(fishing_effort_by_type_05$hours_sq_km)
fishing_effort_by_type_05$logHours = ifelse(fishing_effort_by_type_05$logHours >= 5, 
                                            5, 
                                            fishing_effort_by_type_05$logHours)
fishing_effort_by_type_05$logHours = ifelse(fishing_effort_by_type_05$logHours <= 1, 
                                            1, 
                                            fishing_effort_by_type_05$logHours)

purse_seine_effort_05 = fishing_effort_by_type_05 %>% 
    filter(geartype == 'purse_seines'& 
               lat_bin > -90 & 
               lat_bin < 90 & 
               lon_bin > -180 & 
               lon_bin < 180)


longline_effort_05 = fishing_effort_by_type_05 %>%
    filter(geartype == 'drifting_longlines'& 
               lat_bin > -90 & 
               lat_bin < 85 & 
               lon_bin > -180 & 
               lon_bin < 180)

trawler_effort_05 = fishing_effort_by_type_05 %>%
    filter(geartype == 'trawlers' & 
               lat_bin > -90 & 
               lat_bin < 90 & 
               lon_bin > -180 & 
               lon_bin < 180) 

```


#### Import and process NPP data
```{r, eval=FALSE}
#Generating the Global Raster NPP Figure in R

setwd("../data")
#RcppCNPy allows loading of numpy array into R
library(RcppCNPy)
#load the chlorophyll numpy array and convert to dataframe
#data is 0.25 degree cell npp, so 1440 columns, 720 rows.
chloro = npyLoad("chlorophyll.npy")
chloro = data.frame(chloro)

#assign lat and lon values to row/columns
names(chloro) = as.character(seq(-180,179.75, 0.25))
rownames(chloro) = as.character(seq(-89.75,90, 0.25))

library(dplyr)
#add specify lat as row names using dplyr
chloro = add_rownames(chloro, var = "lat")

library(tidyr)
#reshape into 'long' format...row names are already included in gather
#so just have to specify lon, and assign values to npp
chloro.2 = gather(chloro, "lon", "npp", 2:ncol(chloro))
chloro.2 = data.frame(chloro.2)
chloro.2$lat = as.numeric(chloro.2$lat)
chloro.2$lon = as.numeric(chloro.2$lon)
chloro.2$npp = as.numeric(chloro.2$npp)
chloro.2$npp = ifelse(chloro.2$npp == 0, NA, chloro.2$npp)
chloro.2$npp = ifelse(chloro.2$npp > 20000, 20000, chloro.2$npp)

library(RColorBrewer)
```


#### Generating Fishing Effort Figure Panels
```{r, fig.height=8, fig.width=15}

cont_fill = 'white'#'grey30' #'grey80'
back_fill =  'white'#'#202A50' #'black' #'#1F2A4F'
cont_color = 'grey60'#'grey30'
#color_grad = c( "#414487", "#2A788E", "#22A884", "#7AD151","#FDE725","#FFC04C")
color_grad = c("#C6DBEF", "#9ECAE1", "#6BAED6","#4292C6" ,"#2171B5", "#084594") #blues
```
##### Total Effort
```{r}
total_effort_05 <- ggplot() +
    geom_raster(data = total_fishing_effort_05, 
                aes(x = lon_bin, y = lat_bin, fill = logHours), 
                interpolate = FALSE) +
    geom_sf(data = land_df, 
                 color = cont_color,
                 size = 0.5,
                 fill = cont_fill) +
    scale_fill_viridis("Vessel Density", na.value = NA, limits = c(1, 5), 
                       breaks=c(seq(1,5,length.out = 4)),
                       labels = c('1','10', '100','> 1000' ), 
                       direction = -1)+
    coord_sf(datum = NA) +
    theme_gfw_paper +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.background = element_rect(color = NA,fill = back_fill),
          legend.position = 'none',
          legend.direction="horizontal",
          plot.margin=unit(c(0,-0.5,-0.5,-0.5), 'cm'),
          legend.text = element_text(size = 6, family = 'Times New Roman'),
          legend.title = element_text(size = 7, family = 'Times New Roman'))
total_effort_05
```
  
##### Purse Seine Effort  
  
```{r}
purse_seine_05 <- ggplot() +
    geom_raster(data = purse_seine_effort_05, 
                aes(x = lon_bin, y = lat_bin, fill = logHours), 
                interpolate = TRUE) +
    geom_sf(data = land_df, 
                 color = cont_color,
                 size = 0.5,
                 fill = cont_fill) +
   scale_fill_viridis("Vessel Density", na.value = NA, limits = c(1, 5), 
                       breaks=c(seq(1,5,length.out = 4)),
                       labels = c('1','10', '100','> 1000' ), 
                       direction = -1) +
    coord_sf(datum = NA) +
    theme_gfw_paper +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.border = element_blank(),
          panel.background = element_rect(color = NA,fill = back_fill),
          legend.position = 'none',
          plot.margin=unit(c(0,-0.5,-0.5,-0.5), 'cm'))
purse_seine_05
```
  
##### Drifting Longline Effort  
  
```{r}

longline_05 <- ggplot() +
    geom_raster(data = longline_effort_05, 
                aes(x = lon_bin, y = lat_bin, fill = logHours), 
                interpolate = TRUE) +
    geom_sf(data = land_df,
                 color = cont_color,
                 size = 0.5,
                 fill = cont_fill) +
    scale_fill_viridis("Vessel Density", na.value = NA, limits = c(1, 5), 
                       breaks=c(seq(1,5,length.out = 4)),
                       labels = c('1','10', '100','> 1000' ), 
                       direction = -1) +
    coord_sf(datum = NA) +
    theme_gfw_paper +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.background = element_rect(color = NA,fill = back_fill),
          legend.position = 'none',
          plot.margin=unit(c(0,-0.5,-0.5,-0.5), "cm"))
longline_05
```
  
##### Trawler Effort  
  
```{r}
trawling_05 <- ggplot() +
    geom_raster(data = trawler_effort_05, 
                aes(x = lon_bin, y = lat_bin, fill = logHours), 
                interpolate = TRUE) +
    geom_sf(data = land_df, 
                 color = cont_color,
                 size = 0.5,
                 fill = cont_fill) +
    scale_fill_viridis("Vessel Density", na.value = NA, limits = c(1, 5), 
                       breaks=c(seq(1,5,length.out = 4)),
                       labels = c('1','10', '100','> 1000' ), 
                       direction = -1) +
    coord_sf(datum = NA) +
    theme_gfw_paper +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.background = element_rect(color = NA,fill = back_fill),
          legend.position = 'none',
          plot.margin=unit(c(-0.5,-0.5,-0.5,-0.5), "cm"))
trawling_05
```

#### Generating NPP panel
```{r, eval=FALSE}
color_grad = (c('white',"#EDF8E9", "#C7E9C0", "#A1D99B", "#74C476", "#31A354", "#006D2C")) #white -> greens 
cont_fill = 'white'#'grey30'
back_fill = 'white'#'black' #'#202A50'

chloro.2 <- chloro.2 %>% 
    mutate(logNPP = log10(npp)) %>% 
    mutate(logNPP = ifelse(logNPP >= 4.25, 4.25, logNPP)) %>%
    mutate(logNPP = ifelse(logNPP <= 3.25, 3.25, logNPP))

cont_fill = 'white'
npp_plot = ggplot() +
    geom_raster(data = chloro.2, aes(lon, lat, fill = npp)) +
    geom_polygon(data = land_df, aes(long, lat, group = group), 
                 color = 'grey60', fill = cont_fill ) +
    scale_fill_gradient2('Net Primary Productivity (x1000)',  
                         low = 'white',#e7f9dc',
                         mid = '#f7fcf4', #e6f9da',
                         high = '#31a354',
                         midpoint = 4000,
                         na.value='white',
                         breaks = c(seq(0,20000, 5000)),
                         limits = c(0,20000),
                         labels = c('0', '5','10','15','20'),
                         guide = 'none') +
    theme_gfw_paper +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.grid = element_blank(),
          panel.background = element_rect(fill = 'white'),
          plot.margin=unit(c(0,-0.5,-0.5,-0.5), "cm"))


npp_plot
```

#### Generating Legends
Code for generating legend to a a set of gridded fishing plots. 
Requires making a plot with a legend and then extracting the legend using get_legend
First for the fishing effort and the for the NPP

```{r, eval=FALSE}
library(cowplot)
#plot with legend
#color_grad = c( "#414487", "#2A788E", "#22A884", "#7AD151","#FDE725","#FFC04C")

legend_plot_05 = ggplot() +
    geom_raster(data = total_fishing_effort_05, 
                aes(x = lon_bin, y = lat_bin, fill = logHours), 
                interpolate = FALSE) +
    geom_polygon(data = land_df,
                 fill = cont_fill) +
    scale_fill_viridis("Vessel Density", 
                       na.value = NA, 
                       limits = c(1, 5),
                       breaks=c(seq(1,5,length.out = 5)),
                       direction = -1) +
    guides(fill = guide_colorbar(barwidth = 12, 
                                 barheight = 0.3, 
                                 title.hjust = 0.5,  
                                 title.position = 'bottom'))

effort_legend_05 <- get_legend(legend_plot_05 + 
                                   theme(legend.position="bottom"))


```


Generating NPP legend
```{r, eval=FALSE}

cont_fill = 'white'
npp_legend_plot = ggplot() +
    geom_raster(data = chloro.2, 
                aes(lon, lat, fill = npp)) +
    geom_polygon(data = land_df, 
                 fill = cont_fill ) +
    scale_fill_gradient2('Net Primary Productivity (x1000)', 
                        low = 'white',#e7f9dc', 
                        mid = '#f7fcf4', #e6f9da', 
                        high = '#31a354', 
                        midpoint = 4000, na.value="white",
                        breaks = c(seq(0,20000, 5000)),
                        limits = c(0,20000),
                        labels = c('0', '5','10','15','20')) +
    guides(fill = guide_colorbar(barwidth = 16, 
                                 barheight = 2, 
                                 title.position = 'bottom',
                                 title.hjust = 0.7)) +
    theme_gfw_paper +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.grid = element_blank(),
          #legend.position = c(0.7, .09),
          legend.direction="horizontal",
          legend.text = element_text(size = 14, family = 'Times New Roman'),
          legend.title = element_text(size = 16, family = 'Times New Roman'),
          plot.margin=unit(c(-0.5,0,-0.5,-0.5), 'cm'),
          legend.position = 'bottom')

npp_legend <- get_legend(npp_legend_plot + theme(legend.position="bottom"))

```
